---
title: "What Makes a Highly-Rated Cup of Coffee?"
description: |
    Investigating which characteristics of coffee correlated most strongly with overall coffee ratings.
author: Celina
date: September 30, 2026
format: html
execute:
  warning: false
  message: false
---
## Coffee Ratings
This analysis examines coffee ratings based on data from TidyTuesday (2020-07-07). This is an older dataset that contains variables for various coffee characteristics, and in my analysis I will investigate which of these are most strongly correlated with a coffee's overall rating.

#### Load Data
```{r}
coffee_ratings <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2020/2020-07-07/coffee_ratings.csv')
```

#### Analysis
```{r}
coffee_characteristics <- coffee_ratings |>
  dplyr::select(
    total_cup_points,
    aroma,
    flavor,
    aftertaste,
    acidity,
    body,
    balance,
    uniformity,
    clean_cup,
    sweetness,
    cupper_points
  )


correlations <- coffee_characteristics |>
  cor(use = "complete.obs") |>
  as.data.frame() |>
  dplyr::select(total_cup_points) |>
  dplyr::rename(correlation = total_cup_points) |>
  tibble::rownames_to_column("characteristic") |>
  dplyr::filter(characteristic != "total_cup_points") |>
  dplyr::arrange(correlation)

```

#### Plot

```{r}
library(ggplot2)

ggplot(
  correlations,
  aes(
    x = correlation,
    y = reorder(characteristic, correlation)
  )
) +
  geom_col(fill = "purple") +
  labs(
    title = "Which Characteristics Relate Most to Overall Coffee Rating (cupper points)?",
    subtitle = "Correlation with total cupper points",
    x = "Correlation with total cupper points",
    y = "Coffee characteristic"
  ) +
  theme_minimal()
```


## Complete Citations
[TidyTuesday resource](https://github.com/rfordatascience/tidytuesday/tree/main/data/2020/2020-07-07)
The data was uploaded to the TidyTuesday repository by Jon Harmon. 

[Original data source](https://github.com/jldbc/coffee-quality-database)
The original data can be found on James LeDoux's github. LeDoux is a Buzzfeed Data Scientist who created this database based on reviews of arabica and robusta coffee beans from the Coffee Quality Institute's trained reviewers.
